### 2024-12-02 docs-site/index.Rmd
add docs

% Sobre

Documentação do projeto de disciplina de Ciência de Dados Geoespacial - CAP-423.

## Grupo

Professores:

- Dr. Gilberto Ribeiro
- Dra. Lubia Vinha
- Dr. Rafael Santos
- Dr. Thales Körting

Membros:

- André
- Antonio
- Daniel
- Felipe

## Metodologia

Considerations about sampling, study design, randomization, factors, response variables, etc.