### 2024-12-09 docs-site/code.Rmd
update code

% Code

## Leitura dos dados produzidos em R

## Carregand

The packages necessary for this analysis..

```{r}
library(readr)
```

## Data import

Let's first import the data.

```{r message=FALSE, warning=FALSE}
data <- read_delim("data/data.csv", delim = ";")
data
```

## Data tidy

Data wrangling here.

## Data transformation

We need to do some transformations, create new variables, etc.

## Data visualization

Produce some plots to understand the data

## Data analysis

Modeling

## Manuscript-ready plots

This could be shown here or in a separate webpage.